---
title: "Broad-sense heritability of each trait"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=F)
options(width = 300)
library(knitr)        # CRAN v1.26
library(reshape2)     # CRAN v1.4.4
library(dplyr)        # CRAN v1.0.0
library(tidyverse)    # CRAN v1.3.0
library(kableExtra)   # CRAN v1.1.0
library(rstan)        # CRAN v2.19.2
library(broom)        # CRAN v0.5.2
library(loo)          # CRAN v2.2.0
library(latex2exp)    # CRAN v0.4.0
library(parallel)
library(tidytidbits)  # CRAN v0.2.3
library(tidybayes)    # CRAN v2.0.1
library(bayesplot)    # CRAN v1.7.1
library(cowplot)      # CRAN v1.0.0
library(xtable)
theme_set(theme_bw(base_size = 20))
color_scheme_set("green")
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


```{r OptionDocs}
# Options for sampling
n_chains <- 4
n_iter <- 2500
n_warm <- 1250
n_thin <- 1
save_warmup = FALSE 

theme_set(theme_bw(base_size = 20))

# option for the MCMC intervals
point.est="median"

conf.level <- 0.95
```

# Model equation and code

In this document, we estimate the broad-sense heritability of each trait $y_{bpcr}$ based on the following quantitative genetic model:

\begin{equation}
\begin{aligned}
y_{bpcr} & \sim  \mathcal{N}(\mu_{bpc},\sigma^{2}_{r})\\[3pt]
\mu_{bpc} & =  \beta_{0} +  B_{b} + P_{p} + C_{c(p)}\\[3pt]
\end{aligned}
\end{equation}

where $\beta_{0}$ is the global intercept, $B_{b}$ the block intercepts, $P_{p}$ the population intercepts, $C_{c(p)}$ the clone intercepts and $\sigma^{2}_{r}$ the residual variance.

The prior of $\beta_{0}$ was weakly informative and centered around the mean of the observed values for the trait under considered, as follows:

$$\beta_{0} \sim \mathcal{N}(\mu_{y},2)$$

The population, clone and block intercepts, $P_{p}$, $C_{c(p)}$ and $B_{b}$ were considered normally-distributed with variances $\sigma^{2}_{P}$, $\sigma^{2}_{C}$ and $\sigma^{2}_{B}$, such as:

$$\begin{bmatrix} B_{b} \\ P_{p}\\ C_{c(p)} \end{bmatrix} 
    \sim \mathcal{N}\left(0,
\begin{bmatrix}\sigma^{2}_{B}\\[3pt]
\sigma^{2}_{P}\\
\sigma^{2}_{C}\\
\end{bmatrix}
\right)\\[3pt]$$


To partition the total variance, we parameterize our model so that only the total variance, $\sigma_{tot}^{2}$ has a prior, such that:

\begin{equation}
\begin{aligned}
\sigma_{tot}^{2} & = \sigma_{r}^{2} + \sigma_{B}^{2} + \sigma_{C}^{2} + \sigma_{P}^{2}\\[3pt]
\sigma_{r} & = \sigma_{tot} \times \sqrt(\pi_{r})\\[3pt]
\sigma_{B} & = \sigma_{tot} \times \sqrt(\pi_{B})\\[3pt]
\sigma_{P} & = \sigma_{tot} \times \sqrt(\pi_{P})\\[3pt]
\sigma_{C} & = \sigma_{tot} \times \sqrt(\pi_{C})\\[3pt]
\sigma_{tot} & \sim \mathcal{S}^{*}(0,1,3)
\end{aligned}
\end{equation}

where $\sum_{l}^{4}\pi_{l}=1$ (using the `simplex` function in `Stan`).

Here is the stan model code:

```{r ModelCompilation}
model = stan_model("scripts/StanModels/HierarchicalModel_StandardQuantitativeGeneticModel.stan")
model
```


# Running and saving the models

> Function to download and format the data

```{r FunctionDownloadFormatData,eval=F}
DownloadData <- function(trait,sub){

#### Trait transformations
    # Phenology-realted traits and delta13C are centered to help convergence
if(trait=="MeanBB"|trait=="MeanDBB"|trait=="d13C"){
    CenterResponseVariable <- TRUE
    LogResponseVariable <- FALSE
    
    # SLA is not centered and we used its logarithm as response variable
  }  else if(trait=="POR_SLA"){ 
    CenterResponseVariable <- FALSE
    LogResponseVariable <- TRUE
    
    # Height is not transformed
  } else { 
    CenterResponseVariable <- FALSE
    LogResponseVariable <- FALSE
  }


#### Merging with the phenotypic data
  if(trait=="MeanBB"|trait=="MeanDBB"){
      
  df <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_AnnualTraits.rds") %>% 
    dplyr::filter(!(prov=="MAD")) %>% 
    when(trait=="MeanBB" ~ dplyr::select(.,prov,clon,tree,block,BDX_s32013,BDX_s32014,BDX_s32015,BDX_s32017),
         ~ dplyr::select(.,prov,clon,tree,block,BDX_dbb2014,BDX_dbb2015,BDX_dbb2017)) %>% 
    filter(rowSums(is.na(across(where(is.numeric)))) != ncol(.)-4) %>% 
    dplyr::mutate(trait = rowMeans(select(.,contains("BDX")),na.rm=T)) %>%
    execute_if(CenterResponseVariable, mutate(trait=trait-mean(trait))) %>% 
    execute_if(LogResponseVariable, mutate(trait=log(trait))) %>% 
    arrange(by=tree)
  
  } else {
    
  df <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_AnnualTraits.rds") %>% 
    dplyr::filter(!(prov=="MAD")) %>% 
    dplyr::rename(trait=all_of(trait)) %>% 
    dplyr::select(prov,clon,tree,block,trait) %>% 
    drop_na() %>%  
    execute_if(CenterResponseVariable, mutate(trait=trait-mean(trait))) %>% 
    execute_if(LogResponseVariable, mutate(trait=log(trait))) %>% 
    arrange(by=tree)
  
  }
  
  
########################################################################################################################
# Option to keep only populations represented by at least 4 clones & clones represented by at least 4 individuals.
  if(sub=="sub1"|sub=="sub2"){

  ClonesToKeep <- df %>% 
    dplyr::select(prov,clon) %>% 
    group_by(clon) %>% 
    mutate(NbInd = n()) %>% 
    distinct() %>%
    filter(NbInd>sub.ind)
  
  df <- df %>% semi_join(ClonesToKeep, by="clon") 
  
  PopsToKeep <- df %>%  
    dplyr::select(prov,clon) %>% 
    distinct() %>%  
    group_by(prov) %>%  
    dplyr::summarise(NbClones=n()) %>% 
    filter(NbClones>sub.clones)
  
  df <- df %>% semi_join(PopsToKeep, by="prov") 
  }
########################################################################################################################
    

# the list for stan:

  list.stan <- list(N=length(df$tree),
         y=df$trait,
         nprov = length(unique(df$prov)),
         nclon = length(unique(df$clon)),
         nblock = length(unique(df$block)),
         prov = as.numeric(as.factor(df$prov)),
         clon = as.numeric(as.factor(df$clon)),
         bloc = as.numeric(as.factor(df$block)))
  
}
```

```{r RunModel,eval=F}
########################################
# Running the models on a filtered dataset or not? (modification after the first round of review in Heredity)
sub = "no" # if the dataset is not filtered.

### Two filtered datasets:

# Subset 1: only populations represented by at least 4 clones and clones represented by at least 4 individuals.
sub = "sub1"
sub.clones = 3
sub.ind = 3

# Subset 1: only populations represented by at least 4 clones and clones represented by at least 4 individuals.
sub = "sub2"
sub.clones = 10
sub.ind = 6
########################################

if(sub=="sub2"){traits <- c("BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB")} else {
  traits <- c("POR_htoct12","BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")}


list.models <- list()


for(trait in traits){
  
list.trait <- DownloadData(trait,sub)


list.models[[trait]] <- sampling(model, 
                               data = list.trait, 
                               pars=c("beta0",
                                      "pi","sigma2_r","sigma2_block","sigma2_prov","sigma2_clon",
                                      "H2",
                                      "sigma_tot","sigma_block","sigma_prov","sigma_clon",
                                      "R_squared"),
                               iter = n_iter, 
                               chains = n_chains, 
                               cores = n_chains,
                               save_warmup = save_warmup,
                               thin=n_thin,
                               control=list(max_treedepth=12))
}

if(sub=="sub1"){
saveRDS(list.models,file=paste0("outputs/models/StandardQuantGenModel_sub1.rds"))} else if (sub=="sub2"){
saveRDS(list.models,file=paste0("outputs/models/StandardQuantGenModel_sub2.rds"))} else {
saveRDS(list.models,file=paste0("outputs/models/StandardQuantGenModel.rds"))}
}
```


```{r StandardQuantGenModelH2,fig.height=6,fig.width=8}
sub="no"
# I only made the script for sub="no"
if(sub=="sub2"){traits <- c("BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB")} else {
  traits <- c("POR_htoct12","BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")}

post <- readRDS(file=paste0("outputs/models/StandardQuantGenModel.rds"))  %>% 
  mclapply(function(x) broom::tidyMCMC(x,pars=("H2"),
              droppars = NULL, 
              estimate.method = point.est, 
              ess = F, 
              rhat = F, 
              conf.int = T,
              conf.level = conf.level)) %>% 
  bind_rows() %>% 
  mutate(Traits=traits)

# Table for the Supplementary Information:
post %>% dplyr::select(Traits,estimate,conf.low,conf.high) %>% kable(digits=3,"latex")

traits.order.fig <- c("BDX_htnov18","BDX_htnov13","POR_htoct12","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")
post$Traits <- as.factor(post$Traits)
post$Traits <- factor(post$Traits,levels=rev(c(traits.order.fig)))

d13c <- parse(text = TeX("$\\delta^{13}C$"))

p <- post  %>%   ggplot(aes(x=term,y = estimate,ymin = conf.low, ymax = conf.high,colour=Traits,shape=Traits)) +
  geom_pointinterval(position = position_dodge(width = .6),point_size=2.5,size=3,show.legend = c(size = TRUE)) +
  ylab(TeX("$H^{2}$ estimates")) + xlab("") +
  scale_colour_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)), #
                     values=rev(c("#B2182B","#D73027","#F46D43", "#FDAE61","#5AAE61","#A6DBA0" ,"#4575B4", "#ABD9E9"))) +
  scale_shape_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)),
                     values = c(15, 15,17, 17,16,16,16,16)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.text=element_text(size=10),
      axis.ticks.x = element_blank(),
      legend.background = element_rect(colour = "grey"),
      legend.position = c(.18, .77),
      strip.text.x = element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.x=element_blank()) +
    guides(color=guide_legend(ncol=1,override.aes = list(size=2)))

p

ggsave(p,file="figs/PosteriorDistributions/StandardQuantGenModel_BroadSenseHeritability.png",height=6,width=8)
```



```{r StandardQuantGenModelH2V2,fig.height=6,fig.width=8}
sub="no"
# I only made the script for sub="no"
if(sub=="sub2"){traits <- c("BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB")} else {
  traits <- c("POR_htoct12","BDX_htnov13","BDX_htnov18","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")}

post <- readRDS(file=paste0("outputs/models/StandardQuantGenModel.rds"))  %>% 
  mclapply(function(x) broom::tidyMCMC(x,pars=("H2"),
              droppars = NULL, 
              estimate.method = point.est, 
              ess = F, 
              rhat = F, 
              conf.int = T,
              conf.level = conf.level)) %>% 
  bind_rows() %>% 
  mutate(Traits=traits)

# Table for the Supplementary Information:
post %>% dplyr::select(Traits,estimate,conf.low,conf.high) %>% kable(digits=3,"latex")

traits.order.fig <- c("BDX_htnov18","BDX_htnov13","POR_htoct12","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")
post$Traits <- as.factor(post$Traits)
post$Traits <- factor(post$Traits,levels=rev(c(traits.order.fig)))


p <- post  %>%   
  ggplot(aes(x=estimate,y = Traits,colour=Traits,shape=Traits)) +
  geom_point(size=4) + # 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),height=0.1,size=1,show.legend = c(size = TRUE)) +
  xlab(TeX("$H^{2}$ estimates")) + ylab("") +
  
  scale_y_discrete(labels=c("BDX_htnov18"="Height (Bordeaux, 85 months)",
                            "BDX_htnov13"="Height (Bordeaux, 25 months)",
                            "POR_htoct12"="Height (Portugal, 20 months)",
                            "AST_htnov12"="Height (Asturias, 21 months)",
                            "MeanBB"="Mean bud burst date (Bordeaux)",
                            "MeanDBB"="Mean duration of bud burst (Bordeaux)",
                            "POR_SLA"="Specific Leaf Area (Portugal)",
                            "d13C"=parse(text = TeX("$\\delta^{13}C$ (Portugal)"))
                            )) +
  
  
  scale_colour_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)), #
                     values=rev(c("#B2182B","#D73027","#F46D43", "#FDAE61","#5AAE61","#A6DBA0" ,"#4575B4", "#ABD9E9"))) +
  scale_shape_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)),
                     values = c(15, 15,17, 17,16,16,16,16)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.text=element_text(size=10),
      axis.ticks.x = element_blank(),
      legend.position = "none",
      strip.text.x = element_blank(),
      panel.grid.minor.y=element_blank(),
      panel.grid.major.y=element_blank()) +
    guides(color=guide_legend(ncol=1,override.aes = list(size=2)))

p

ggsave(p,file="figs/PosteriorDistributions/StandardQuantGenModel_BroadSenseHeritability_V2.png",height=8,width=14)
```



```{r StandardQuantGenModelPI,fig.height=8,fig.width=12}
post <- readRDS(file=paste0("outputs/models/StandardQuantGenModel.rds"))  %>% 
  mclapply(function(x) broom::tidyMCMC(x,pars=("pi"),
              droppars = NULL, 
              estimate.method = point.est, 
              ess = F, 
              rhat = F, 
              conf.int = T,
              conf.level = conf.level) %>% dplyr::mutate(pi=c("Residuals","Blocks","Populations","Clones"))) %>% 
  bind_rows() %>% 
  mutate(Traits=sapply(traits,function(x) rep(x,4),simplify = T) %>% as.vector())

traits.order.fig <- c("BDX_htnov18","BDX_htnov13","POR_htoct12","AST_htnov12","MeanBB","MeanDBB","POR_SLA","d13C")
post$Traits <- as.factor(post$Traits)
post$Traits <- factor(post$Traits,levels=rev(c(traits.order.fig)))

d13c <- parse(text = TeX("$\\delta^{13}C$"))

p <- post  %>%   ggplot(aes(x = pi, y = estimate,ymin = conf.low, ymax = conf.high,colour=Traits,shape=Traits)) +
  geom_pointinterval(position = position_dodge(width = .6),point_size=2.5,size=3,show.legend = c(size = TRUE)) +
  ylab(TeX("Proportion of variance explained")) + xlab("") +
  scale_colour_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)), # 
                     values=rev(c("#B2182B","#D73027","#F46D43", "#FDAE61","#5AAE61","#A6DBA0" ,"#4575B4", "#ABD9E9"))) +
  scale_shape_manual(labels = rev(c("Height (Bordeaux, 85 months)",
                                "Height (Bordeaux, 25 months)",
                                "Height (Portugal, 20 months)",
                                "Height (Asturias, 21 months)",
                                "Mean bud burst date",
                                "Mean duration of bud burst",
                                "Specific Leaf Area",
                                d13c)),
                     values = c(15, 15,17, 17,16,16,16,16)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=13),
        axis.text.y = element_text(size=16),
        axis.title = element_text(size=16),
      legend.title=element_text(size=13), 
      legend.position = c(0.22,0.88),
      legend.text=element_text(size=10),
      legend.background = element_rect(colour = "grey"),
      strip.text.x = element_text(size = 14),
      panel.grid.minor.x=element_blank(),
      panel.grid.major.x=element_blank()) +
    guides(color=guide_legend(ncol=2,override.aes = list(size=2)))
p

ggsave(p,file="figs/PosteriorDistributions/StandardQuantGenModel_VariancePartitioning.png",height=8,width=12)
```

