---
title: "Mapping - Figure 1"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
# knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=TRUE,cache.lazy = FALSE)
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=F)
options(width = 300)
library(knitr)        # CRAN v1.26
library(reshape2)     # CRAN v1.4.4
library(dplyr)        # CRAN v1.0.0
library(tidyverse)    # CRAN v1.3.0
library(kableExtra)   # CRAN v1.1.0
library(cowplot)      # CRAN v1.0.0
library(parallel)
library(sf)           # CRAN v0.9-4
library(readxl)
library(raster)       # CRAN v3.3-13
library(spData)       # CRAN v0.3.5
library(RStoolbox)    # CRAN v0.2.6
library(rnaturalearth)
```


# Provenance and site coordinates

```{r ProvSiteCoordinates}
# Load CLONAPIN phenotypic data
data <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_AnnualTraits.rds")

# Provenance coordinates
# ----------------------

# >> CLONAPIN provenances (33 provenances)
# -----------------------------------------

df <- unique(data[,c("prov","longitude_prov","latitude_prov")])

# Remove the ROD population because it has no genomic data,
# and the MAD population because it has only one clone (so the genetic variation can't
# be estimated)
df <- df[!(df$prov=="ROD"|df$prov=="MAD"),]
colnames(df) <- c("prov","longitude","latitude")

# Create a spatial object of the provenance coordinates 
xyprov <- SpatialPoints(df[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))

# Reproject in ETRS89 - LAEA 
xyprov <- spTransform(xyprov,crs(raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif"))))

# for mapping
xyprovdf <- xyprov@coords %>% as_tibble()


# >> 23 provenances in common with the progeny test in Asturias
# -------------------------------------------------------------
# Load the progeny test data of Asturias
prog.test.data <- read_excel("data/ProgenyTests/F26SNP_d130809_20120515.xls", sheet = 2) %>% drop_na(Prov_code)

# Load the file with the matching of code between CLONAPIN (column "CODE") and the progeny tests (column "CODE_FIELD")
prov.info <- read_csv(file="data/ClonapinData/coordinates_provenances.csv") %>% 
  filter(CODE %in% unique(df$prov)) %>%  # keep only the 33 provenances of CLONAPIN used in the study
  filter(CODE_FIELD %in% unique(prog.test.data$Prov_code)) # keep only the provenances of the progeny tests

# Keep only the 23 provenances in common
df.common <- df[df$prov %in% prov.info$CODE,]

# Create a spatial object of the provenance coordinates 
xyprov.common <- SpatialPoints(df.common[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))

# Reproject in ETRS89 - LAEA 
xyprov.common <- spTransform(xyprov.common,crs(raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif"))))

# for mapping
xyprovdf.common <- xyprov.common@coords %>% as_tibble()

# >> Provenances in CLONAPIN but not in the progeny test in Asturias
# ------------------------------------------------------------------

# Keep only the 10 provenances that are in CLONAPIN but not in the progeny test
df.diff <- df[!(df$prov %in% prov.info$CODE),]

# Create a spatial object of the provenance coordinates 
xyprov.diff <- SpatialPoints(df.diff[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))

# Reproject in ETRS89 - LAEA 
xyprov.diff <- spTransform(xyprov.diff,crs(raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif"))))

# for mapping
xyprovdf.diff <- xyprov.diff@coords %>% as_tibble()



# ------------------------------------------------------------------------------------------------------------------------------
# Site coordinates
# ----------------

df <- unique(data[,c("site","longitude_site","latitude_site")])
colnames(df) <- c("site","longitude","latitude")

# Create a spatial object of the site coordinates 
xysite <- SpatialPoints(df[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))

# Reproject in ETRS89 - LAEA
xysite <- spTransform(xysite,crs(raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif"))))

# for mapping
xysitedf <- xysite@coords %>% as_tibble() %>% mutate(Site=c("Asturias","Bordeaux","Fundão"))
```



```{r MaritimePineDistri}
# Create a buffer of 10 km around the population and site location
buffer.prov <- buffer(xyprov,width=10000)
buffer.site <- buffer(xysite,width=10000)

# Load the distribution of maritime pine based on Euforgen and NFI plots
PinpinDistri  <- shapefile('../../GenomicOffset/GenomicOffsetPinPin/data/maps/MaskPinpinDistri/PinpinDistriEUforgen_NFIplotsBuffer10km.shp')
PinpinDistri <- spTransform(PinpinDistri,crs(raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif")) ))

# Merge maritime pine distribution and 10-km buffer around the population and site location
PinpinDistri <- rgeos::gUnion(PinpinDistri,buffer.prov)
PinpinDistri <- rgeos::gUnion(PinpinDistri,buffer.site)
```

# PCA of the 2 climatic variables

$EMT$: Extreme minimum temperature over the period considered (1901-1950) (°C)

$SHM$: Summer heat moisture index (°C/mm)

The CRS of the climatic rasters is **EPSG:3035 (ETRS89, LAEA)** (+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs).

```{r PCA,fig.height=5,fig.width=5}
ext <- extent(matrix(c(2500000,1200000,4500000,3000000), nrow=2)) # extent for the map visualization

var.clim <- c("EMT","SHM")

clim.val <- var.clim %>% mclapply(function(x){

  rast <- raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/",x,".tif")) 
  rast <- crop(rast,ext)
  rast <- mask(rast,PinpinDistri,updatevalue=NA)
  val <- getValues(rast)
  val <- tibble(cell.num=1:length(val),"{x}" :=val) %>% drop_na()
  
  })

clim.val <- clim.val[[1]] %>% 
  inner_join(clim.val[[2]],by="cell.num")

# Run the PCA with 2 variables: EMT and SHM
# center and scale the PCA
pca <- prcomp(clim.val[2:3], center = T, scale. = T)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2];
r <- a1; g <- -a2-a1 ;b <- a2 
g <- a1; b <- -a2-a1 ;r <- a2 


# One raster for each color
rast2 <- rast3 <- rast1 <- raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif")) %>% 
  crop(ext) %>% 
  mask(PinpinDistri,updatevalue=NA)

# scale colors and assign color to raster
rast1[clim.val$cell.num] <- (r-min(r))/(max(r)-min(r))*255
rast2[clim.val$cell.num] <- (g-min(g))/(max(g)-min(g))*255
rast3[clim.val$cell.num] <- (b-min(b))/(max(b)-min(b))*255

# Stack together the three rasters (one for each color)
outRast <- stack(rast1, rast2, rast3)

# visualize
plotRGB(outRast, r = 1, g = 2, b = 3, bgalpha=0)
```


# Mapping for Figure 1

```{r MapPCA,fig.height=10,fig.width=10}
world <- ne_countries(scale = "medium", returnclass = "sf")

p <- ggRGB(outRast, r=1,g=2, b=3) + 
  geom_sf(data = world,color=alpha("gray",0.8),size=0.2,fill=alpha("gray",0.1))  +
    coord_sf(xlim = c(2500000, 4500000), 
             ylim = c(1200000,3000000),
             crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ") + 
  geom_point(data=xyprovdf,aes(x=longitude,y=latitude),size=2) +
  geom_point(data=xysitedf,aes(x=longitude,y=latitude),col="red",size=3,shape=17) +
  geom_text(data=xysitedf,aes(x=longitude,y=latitude,label=Site),hjust=1.2, vjust=-1.2, col="red",size=6,fontface="bold") +
  theme_map()
p
#ggsave(p,file="maps/maprgb.svg") # then I used Inkscape to do the final map

p <- ggRGB(outRast, r=1,g=2, b=3) + 
  geom_sf(data = world,color=alpha("gray",0.8),size=0.2,fill=alpha("gray",0.1))  +
    coord_sf(xlim = c(2500000, 4500000), 
             ylim = c(1200000,3000000),
             crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ") + 
  geom_point(data=xyprovdf.diff,aes(x=longitude,y=latitude),size=2) +
  geom_point(data=xyprovdf.common,aes(x=longitude,y=latitude),size=2,shape=9) +
  geom_point(data=xysitedf,aes(x=longitude,y=latitude),col="red",size=3,shape=17) +
  geom_text(data=xysitedf,aes(x=longitude,y=latitude,label=Site),hjust=1.2, vjust=-1.2, col="red",size=6,fontface="bold") +
  theme_map()
p

#ggsave(p,file="maps/maprgb_progtest.svg") # then I used Inkscape to do the final map
```


```{r PlottingPCA,eval=T,fig.height=5,fig.width=5}
nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 0.5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1

# put the PC1 and PC2 scores in the same dataset as the one where there are the cell values
clim.val$PCA1 <- pca$x[, 1]
clim.val$PCA2 <- pca$x[, 2]

cell.num.prov <- raster::extract(rast1,xyprov,cellnumbers=T)[,"cells"]
cell.num.prov.common <- raster::extract(rast1,xyprov.common,cellnumbers=T)[,"cells"]
cell.num.prov.diff <- raster::extract(rast1,xyprov.diff,cellnumbers=T)[,"cells"]
cell.num.site <- raster::extract(rast1,xysite,cellnumbers=T)[,"cells"]


# Option with all provenances represented in the same way
#png("maps/pca.png", width = 1000, height = 1000)
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(rast1[clim.val$cell.num] , 
                                                                             rast2[clim.val$cell.num], 
                                                                             rast3[clim.val$cell.num], max = 255), asp = 1)
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625,lwd=3)
jit <- 0.6
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec,cex=3)
points(clim.val$PCA1[clim.val$cell.num %in% cell.num.prov],
       clim.val$PCA2[clim.val$cell.num %in% cell.num.prov], pch = 16,cex=1.5)

points(clim.val$PCA1[clim.val$cell.num %in% cell.num.site],
       clim.val$PCA2[clim.val$cell.num %in% cell.num.site], pch = 17,cex=1.5,col="red")
#dev.off() 


# Option that differentiate provenance in common or not with the progeny test data
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(rast1[clim.val$cell.num] , 
                                                                             rast2[clim.val$cell.num], 
                                                                             rast3[clim.val$cell.num], max = 255), asp = 1)
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625,lwd=3)
jit <- 0.6
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec,cex=3)
points(clim.val$PCA1[clim.val$cell.num %in% cell.num.prov.common],
       clim.val$PCA2[clim.val$cell.num %in% cell.num.prov.common], pch = 9,cex=1.5)
points(clim.val$PCA1[clim.val$cell.num %in% cell.num.prov.diff],
       clim.val$PCA2[clim.val$cell.num %in% cell.num.prov.diff], pch = 16,cex=1.5)
points(clim.val$PCA1[clim.val$cell.num %in% cell.num.site],
       clim.val$PCA2[clim.val$cell.num %in% cell.num.site], pch = 17,cex=1.5,col="red")
```


