---
title: "Mapping - Figure 1"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
# knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=TRUE,cache.lazy = FALSE)
knitr::opts_chunk$set(fig.width = 5,fig.height = 4,cache=F)
options(width = 300)
library(knitr)        # CRAN v1.26
library(reshape2)     # CRAN v1.4.4
library(dplyr)        # CRAN v1.0.0
library(tidyverse)    # CRAN v1.3.0
library(kableExtra)   # CRAN v1.1.0
library(cowplot)      # CRAN v1.0.0
library(parallel)
library(sf)           # CRAN v0.9-4
library(raster)       # CRAN v3.3-13
library(spData)       # CRAN v0.3.5
library(RStoolbox)    # CRAN v0.2.6
```


# Provenance and site coordinates

```{r ProvSiteCoordinates}
data <- readRDS(file="data/ClonapinData/PhenoDataNovember2019_AnnualTraits.rds")

# Provenance coordinates
# ----------------------
df <- unique(data[,c("prov","longitude_prov","latitude_prov")])

# Remove the ROD population because it has no genomic data,
# and the MAD population because it has only one clone (so the genetic variation can't
# be estimated)
df <- df[!(df$prov=="ROD"|df$prov=="MAD"),]
colnames(df) <- c("prov","longitude","latitude")

                      # Create a spatial object of the provenance coordinates 
xyprov <- SpatialPoints(df[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))


# Site coordinates
# ----------------
df <- unique(data[,c("site","longitude_site","latitude_site")])
colnames(df) <- c("site","longitude","latitude")

                      # Create a spatial object of the site coordinates 
xysite <- SpatialPoints(df[,c("longitude","latitude")], 
                        proj4string=CRS("+proj=longlat +datum=WGS84"))
```

# PCA of the 2 climatic variables

$EMT$: Extreme minimum temperature over the period considered (1901-1950) (°C)

$SHM$: Summer heat moisture index (°C/mm)

The CRS of the climatic rasters is **EPSG:3035 (ETRS89, LAEA)** (+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs).

```{r PCA,fig.height=5,fig.width=5}
ext <- extent(matrix(c(2500000,1200000,4500000,3000000), nrow=2))

var.clim <- c("EMT","SHM")

clim.val <- var.clim %>% mclapply(function(x){

  rast <- raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/",x,".tif")) 
  rast <- crop(rast,ext)
  val <- getValues(rast)
  val <- tibble(cell.num=1:length(val),"{x}" :=val) %>% drop_na()
  
  })

clim.val <- clim.val[[1]] %>% 
  inner_join(clim.val[[2]],by="cell.num")

# Run the PCA with 2 variables: EMT and SHM
pca <- prcomp(clim.val[2:3], center = T, scale. = T)

##assigns to colors, edit as needed to maximize color contrast, etc.
a1 <- pca$x[,1]; a2 <- pca$x[,2];
r <- a1; g <- -a2-a1 ;b <- a2 
g <- a1; b <- -a2-a1 ;r <- a2 


# scale colors and assign color to raster
rast1 <- rast2 <- rast3 <- raster(paste0("data/Climate/Raw/FromMaurizio/Annualndices/EMT.tif")) %>% crop(ext)
rast1[clim.val$cell.num] <- (r-min(r))/(max(r)-min(r))*255
rast2[clim.val$cell.num] <- (g-min(g))/(max(g)-min(g))*255
rast3[clim.val$cell.num] <- (b-min(b))/(max(b)-min(b))*255


outRast <- stack(rast1, rast2, rast3)
plotRGB(outRast, r = 1, g = 2, b = 3,bgalpha=0)
```


```{r SiteProvCoordinatesLAEA}
# Reproject in ETRS89 - LAEA 
xyprov <- spTransform(xyprov,crs(rast1))
xysite <- spTransform(xysite,crs(rast1))

xyprovdf <- xyprov@coords %>% as_tibble()
xysitedf <- xysite@coords %>% as_tibble() %>% mutate(Site=c("Asturias","Bordeaux","Fundão"))
```


# Mapping

```{r MapPCA,fig.height=6,fig.width=6}
p <- ggRGB(outRast, r=1,g=2, b=3) + 
  geom_point(data=xyprovdf,aes(x=longitude,y=latitude),size=2) +
  geom_point(data=xysitedf,aes(x=longitude,y=latitude),col="red",size=3,shape=17) +
  geom_text(data=xysitedf,aes(x=longitude,y=latitude,label=Site),hjust=0, vjust=2, col="red",size=3) +
  theme_map()
#ggsave(p,file="maps/maprgb.svg") # then I used Inkscape to do the final map
p
```


```{r PlottingPCA,eval=F}
nvs <- dim(pca$rotation)[1]
vec <-  row.names(pca$rotation)
lv <- length(vec)
vind <- rownames(pca$rotation) %in% vec
scal <- 0.5
xrng <- range(pca$x[, 1], pca$rotation[, 1]/scal) *1.1
yrng <- range(pca$x[, 2], pca$rotation[, 2]/scal) * 1.1

png("maps/pca.png", width = 500, height = 500)
plot((pca$x[, 1:2]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(scalR , scalG, scalB, max = 255), asp = 1)
points(pca$rotation[!vind, 1:2]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), pca$rotation[vec,1]/scal, pca$rotation[vec, 2]/scal, length = 0.0625,lwd=5)
jit <- 0.6
text(pca$rotation[vec, 1]/scal + jit * sign(pca$rotation[vec, 1]), 
     pca$rotation[vec, 2]/scal + jit * sign(pca$rotation[vec, 2]), 
     labels = vec,cex=3)
dev.off() 
```


